Clause size reduction with all-UIP Learning

### 2020-02-18 paper.tex
address comments and pass over evaluation section, ongoing

\section{Implementation and Experiments}
\subsubsection{Clause Reduction with $\allUip$.}
We implemented our new clause learning schemes in the
\text{\MapleBase} \cite{}, the winner of SAT Race 2015 application
track. We then evaluated our schemes on the full set of benchmarks
from SAT RACE 2019 main track which contains 400 instances. We ran our
experiments on \fb{insert GHz of CPU}, allowing \fb{insert time bound}
seconds per instance and a maximum of \fb{insert RAM bound}.

\fb{Delete figure 2 and use only figure 8 with all of the
  solvers. First briefly comment on the performance and make a brief
  observation about clause sizes.

  Then have a new subsubsection that looks at clause sizes wrt. to the
  two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
  them side by side (using a tabular environment as I did in figure 1).

  Remove fig 5 and put those numbers in the text instead of in a table
  as part of your discussion.

  Then have a new subsubsection on DRAT trim and get rid of Fig 7.

  Then a subsubsection on \allUip in modern solvers. Some of our
  observations have no solid foundation, so check them to make sure
  that you are just not stating your intuitions without reasonable
  evidence. We need to to fit into 15 pages total along with a short
  1/3 page for conclusions.

  Finally add all citations.

  If we add the more complex \allUipMin we also will need more space
  to describe that, so prune some of your conclusions especially those
  that are simply conjectures.
}

Beside solved instances count and PAR-2 score, we additionally measure
the average clause length and clause reduction ratio (both cumulative
and non-cumulative)\footnote{The cumulative reduction ratio is
  obtained through learning all clauses with the target learning
  scheme; Therefore, the reduction is cumulative. The non-cumulative
  reduction ratio is obtained by running the target scheme for
  measurement only (the minimized 1-UIP clause is learned); Therefore,
  the reduction is not cumulative. }  for each instances. For
$\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
learning attempted rate and success rate.

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
\hline
Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
\hline
$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
\hline
$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
\hline
$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 race main track.
CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
\label{fig:t1}
\end{figure}

Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
(five) more instances than the baseline solver with lower PAR-2
scores. $\allUipPure$ has marginally lower PAR-2 score than
$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
respectively. Fig.~\ref{fig:len_pdf} shows the probability density
distribution (PDF) of the relative clause size of both i-uip learning
schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
(88.5\%) of instances, and average relative reduction from 1-UIP is
16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
both i-UIP learning schemes in general produce the smaller clauses,
and the size reduction is more significant for instances with large
average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
green) is consistently smaller than $\allUipPure$ clauses.

2\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF}
    \caption{ Relative clause size reduction distribution. X axis
      indicates the relative size of difference between i-uip and
      1-uip clauses (calculated as
      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
      each instance, and Y axis shows the probability density.}
     \label{fig:len_pdf}
\end{figure}
\begin{figure} \label{fig:len_compare}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/i-uip-sizes-compare.pdf}
    \caption{Average clause size comparison plot. Each point in the
      plot represents an instance. X and Y axis shows the clause
      length from $\allUip$ and 1-UIP, respectively. Each green (red)
      dot represents an compared instance between $\MapleBase$ and
      $\MapleIUIMin$ ($\allUipPure$). }
    \label{fig:len_compare}
\end{figure}
\nf{ Do we need this?  We additionally looked at the 14 instances
  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
  clauses for all of them with average relative reduction of 22\% and
  maximum 77\% (30 vs 135). Seven out of 14 instances has size
  relative reduction over 30\%. For the 9 instances solved by 1-UIP
  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
  three instances and with average relative reduction of 3.3\%.}

$\allUipMin$ outperformed $\allUipPure$ in clause size.  This results agrees with our observation in Fig.~\ref{fig:t2}: $\allUipMin$ attempted $\allUip$ learning more frequently, and it is more likely to succeed.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
\hline
$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
\hline
$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate. $\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to successfully produce smaller $C_i$ clause .}
\label{fig:t2}
\end{figure}

A solver produce smaller clauses can construct smaller proofs. For
UNSAT instances, we measure their DRUP\cite{} proof checking time as
well as the size of the optimized DRUP proof. We used DART-trim
\cite{} with 5000 timeout to check and optimize DRUP proofs.

Fig.~\ref{fig:t3} shows that the optimized proof construct by
$\allUipMin$ and $\allUipPure$ are significantly smaller than 1-UIP
proofs. The relative proof size reduction roughly correlates to the
average clause size reduction. Fig.~\ref{fig:proof_compare} shows the
absolute proof size comparison results.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & optimized proof size (MB) & relative reduction size  \\ 
\hline
$\MapleBase$ & 613.9 & 0  \\ 
\hline
$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
\hline
$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$. Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
\label{fig:t3}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
    \label{fig:proof_compare}
\end{figure}

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
\hline
\nf{$\allUipPure$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
\hline
$\allUipMin$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
\hline
$\allUipAct$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
\hline
$\allUipIn$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
\hline
$\allUipEx$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\allUipPure$, $\allUipMin$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
\label{fig:t4}
\end{figure}

\subsection{$\allUip$ as a Practical Learning Scheme}
To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
benchmark and reported solved instances, PAR-2 score and average
clause size.

Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
scheme $\allUipPure$ solved the most overall instances (228) and the
most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
produced the shortest average clause size, but solved the second least
instances, one more instance than the baseline 1-UIP learning. All
configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
solved instances, PAR-2 score and average clause size.

\subsection{$\allUip$ on Modern SAT solvers}
To validate $\allUip$ as a generalizable learning scheme on modern SAT
solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$). $\expSATShort$
is a top ten solver from 2019 SAT race which uses random walk
simulation to help branching. We chose $\expSATShort$ because 1) it is
a top solver in the 2019 SAT Race without using chronological
backtracking; 2) the combination of random walk simulation and
variable activity branching heuristic allows our learning schemes to
partially sidestep the problem of variable activity. For each solver,
we compare the base 1-UIP learning scheme against $\allUipPure$,
$\allUipMin$ and the top two $\allUip$ variants, $\allUipAct$ and
$\allUipIn$, on the SAT Race 2019 main track benchmark. We report
solved instances, PAR-2 score and the average clause size.

@@

Table~\ref{fig:t5} shows the benchmark result of $\allUip$
configurations on different solvers. All four configurations of
$\allUip$ outperformed 1-UIP learning for the SAT 2017 race winner,
$\MapleSeven$. More specifically, $\allUipPure$, $\allUipMin$,
$\allUipAct$ and $\allUipIn$ solved 12, 8, 5 and 2 more instances,
respectively, whiling producing smaller clauses. $\allUipPure$ solved
more UNSAT and SAT instances while other configurations improved on
solving SAT instances.  The improvement of $\allUip$ is more significant
on $\MapleSeven$ than on $\MapleBase$ for both solved instances and
clause size reduction. This may suggests that $\allUip$ and the recent
learnt clause minimization approach~\cite{} synergies well because
i-UIP clauses are shorter with more common literals which allows
vivification~\cite{} to prune literals more aggressively through unit
propagation.

We observed significant improvement of $\allUip$ for the SAT 2018
winner, $\MapleEightShort$. Three out of four $\allUip$ configurations
outperformed 1-UIP by 5, 4 an 4 instances,
respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using
chronological backtracking (CB) for long distance
backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
configurations. The results indicates that the improvement of $\allUip$
is slightly shadowed by the adoption of CB. More specifically, we
believe CB prevents decision levels from being compressed through the
process of backtracking and literal assertion. The shorter i-UIP
clauses can bring related literals closer, which in turns compresses
the assertion stack and the decision levels. However, since CB
discourages long distance backtracking, the effect of learning shorter
clauses is shadowed until a full restart.  We believe we have observed
an interesting interaction effect that shows the limitations of both
CB and $\allUip$ learning for future research.

$\allUip$ learning schemes showed significant improvement for
$\MapleNineShort$. Three out of four configurations improved solved
instances by 6, 5 and 5 instances, respectively. $\MapleNineShort$
uses duplicate learning (DL) to prioritize clauses that are learned
multiple times. $\allUip$ and DL didn't synergies well possibly because
i-UIP clauses are less likely to be duplicated. We observed that
$\allUipMin$ ,in average, added 12\% less duplicated clauses into the
core clause database than 1-UIP. One possible explanation is that
$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
for different assertion trails. Therefore, the solver is less likely
to learn the same i-UIP clauses.

We observed significant improvement of $\allUip$ for $\expSATShort$,
three out of four configurations of $\allUip$ significantly outperformed
1-UIP learning by 4, 7 and 8 more instances, respectively, while
producing smaller clauses. We expect all three configurations of
$\allUip$ to solve the similar amount of instances with close PAR-2
scores because the additional random walk exploration allows the
learning schemes to partially sidestep the activity problem; hence,
the learning adjustment for variable activities should have less
impact on the solver's performance. However, we instead, observed that
both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
learning scheme. One possible explanation is that $\allUipAct$ and
$\allUipIn$ schemes could easily overcompensate variable activities
for i-UIP clauses, and $\expSATShort$'s random walk exploration could
use future search information to mitigate the negative effects of our
overcompensation.

>>>>

\section{Implementation and Experiments}
%\subsubsection{Clause Reduction with $\allUip$.}
We implemented $\allUip$ learning schemes on
\text{\MapleBase} \cite{liang2016maple}, the winner of SAT Race 2015 application track. We then evaluated these schemes and compare against 1-UIP baseline on the full set of benchmarks from SAT RACE 2019 main track which contains 400 instances. We ran our experiments on 2.70 GHz XeonE5-2680 CPUs 
%\fb{insert GHz of CPU}
, allowing 
%\fb{insert time bound}
5000 seconds per instance and a maximum of 12 GB memory.
%\fb{insert RAM bound}.

%\fb{Delete figure 2 and use only figure 8 with all of the
%  solvers. First briefly comment on the performance and make a brief observation about clause sizes.
 %Then have a new subsubsection that looks at clause sizes wrt. to the
  %two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
  %them side by side (using a tabular environment as I did in figure 1).

  %Remove fig 5 and put those numbers in the text instead of in a table
  %as part of your discussion.

  %Then have a new subsubsection on DRAT trim and get rid of Fig 7.

  %Then a subsubsection on \allUip in modern solvers. Some of our
  %observations have no solid foundation, so check them to make sure
  %that you are just not stating your intuitions without reasonable
  %evidence. We need to to fit into 15 pages total along with a short
  %1/3 page for conclusions.

  %Finally add all citations.

  %If we add the more complex \allUipMin we also will need more space
  %to describe that, so prune some of your conclusions especially those
  %that are simply conjectures.
%}

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
\hline
$\allUipPure$ &\textbf{228} (135, \bf{93}) +7 & 4867.37 & 49.88 \\
\hline
$\allUipMin$ & 226 (135, 91) +5 & 4890.67 & 45.2 \\ 
\hline
$\allUipAct$ & 226 (135, 91)  +5 & \textbf{4866.94} & 47.7 \\
\hline
$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52..12 \\
\hline
$\allUipEx$ & 223 (134, 89) +2& 5015.23& \textbf{43.2} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\allUipPure$, $\allUipMin$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
\label{fig:t4}
\end{figure}

Fig.~\ref{fig:t4} shows each learning scheme's solved instances count, PAR-2 score, and average clause size\footnote{The average clause size is measured by dividing learnt literals by learnt clauses.}.  $\allUip$ learning schemes improved solved instances, and PAR-2 scores over 1-UIP. More specifically, $\allUipPure$ solved the most instances (+7 over 1-UIP) and the most UNSAT instances (+4);  $\allUipIn$ solved the most SAT instances (+6); and $\allUipAct$ yields the best PAR-2 score (-151 than 1-UIP). Besides the performance gain, $\allUip$ schemes yield significantly smaller average clause size.

\subsubsection{Clause Reduction with $\allUip$}
To precisely measure $\allUip$ clause reduction power, we compare each instance's learnt clause size from $\allUipMin$ and $\allUipPure$ against 1-UIP. Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the relative clause size of $\allUip$ learning
schemes ($\allUipMin$ in green and $\allUipPure$ in red) for each instance. $\allUipMin$ ($\allUipPure$ resp.) produces shorter clauses for 88.5\% (77.7\%) of instances, and the average relative reduction rate from 1-UIP is 18.6\% (16.7\%). Fig.~\ref{fig:len_compare} compares the absolute clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.  
Both $\allUip$ schemes generally yield smaller clauses,
and the size reduction is more significant for instances with larger 1-UIP clause size.

Horizontally, The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and \ref{fig:len_compare} indicate $\allUipMin$ often achieves higher clause reduction than $\allUipPure$. We also observed that $\allUipMin$
attempted algorithm~\ref{alg:i-uip} more frequently than $\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed (59.3\% vs 43.4\%). This observation agrees with our experiment results.

\begin{figure}[t!]
{\fss{8pt}{10pt}
\begin{tabular}[t]{l c l}

  \begin{minipage}[t]{0.5\textwidth}
   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF}
   \caption{Relative clause size reduction distribution. X axis
      indicates the relative size of difference between i-uip and
      1-uip clauses (calculated as
      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
      each instance, and Y axis shows the probability density.}
       \label{fig:len_pdf}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/i-uip-sizes-compare.pdf}
    \caption{Average clause size comparison plot. Each point in the
      plot represents an instance. X and Y axis shows the clause
      length from $\allUip$ and 1-UIP, respectively. Each green (red)
      dot represents an compared instance between $\MapleBase$ and
      $\MapleIUIMin$ ($\allUipPure$).}
      \label{fig:len_compare}
  \end{minipage}

\end{tabular}
}
\end{figure}

%Beside solved instances count and PAR-2 score, we additionally measure
%the average clause size. 
%and clause reduction ratio (both cumulative
%and non-cumulative)\footnote{The cumulative reduction ratio is
%  obtained through learning all clauses with the target learning
%  scheme; Therefore, the reduction is cumulative. The %non-cumulative
%  reduction ratio is obtained by running the target scheme for
%  measurement only (the minimized 1-UIP clause is learned); Therefore,
%  the reduction is not cumulative. }  for each instances. 
%For $\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
%learning attempted rate and success rate.

%\begin{figure} 
%\begin{center}
%\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
%\hline
%Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
%\hline
%$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
%\hline
%$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
%\hline
%$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
%\hline
%\end{tabular}
%\end{center}
%\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 %race main track.
%CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, %and the values in the brackets are the non-cumulative reduction ratio.}
%\label{fig:t1}
%\end{figure}

%Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
%(five) more instances than the baseline solver with lower PAR-2
%scores. $\allUipPure$ has marginally lower PAR-2 score than
%$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
%significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
%respectively. Fig.~\ref{fig:len_pdf} shows the probability density
%distribution (PDF) of the relative clause size of both i-uip learning
%schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
%instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
%(88.5\%) of instances, and average relative reduction from 1-UIP is
%16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
%clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
%both i-UIP learning schemes in general produce the smaller clauses,
%and the size reduction is more significant for instances with large
%average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
%5green) is consistently smaller than $\allUipPure$ clauses.

%\nf{ Do we need this?  We additionally looked at the 14 instances
%  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
%  clauses for all of them with average relative reduction of 22\% and
%  maximum 77\% (30 vs 135). Seven out of 14 instances has size
%  relative reduction over 30\%. For the 9 instances solved by 1-UIP
%  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
%  three instances and with average relative reduction of 3.3\%.}

%\begin{figure} 
%\begin{center} 
%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
%\hline
%Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
%\hline
%$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
%\hline
%$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
%\hline
%\end{tabular}
%\end{center}
%\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate. %$\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to %successfully produce smaller $C_i$ clause .}
%\label{fig:t2}
%\end{figure}
\subsubsection{Proof Reduction with $\allUip$}
A learning scheme yields smaller clauses (lemmas) may construct smaller causal proofs. For
UNSAT instances, we additionally compare the size of the optimized DRAT proof from $\allUipPure$, $\allUipMin$ and 1-UIP schemes. We used DART-trim \cite{wetzler2014drat} with 10000 second timeout to check and optimize every DRAT proof once \footnote{Applying DART-trim multiple times can further reduce the proof size until fix-point, but the full optimization is time consuming and impractical for our experiments.}.

The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$ are 413.2MB and 487.2MB, respectively. Both sizes are significantly smaller than the average optimized proof size from 1-UIP, 613.9MB.  The average relative proof size reduction per instance of $\allUipMin$ and $\allUipPure$ are 17.18\% and 6.9\% compare to 1-UIP, which roughly correlate to the observation from relative clause reduction plot in Fig~\ref{fig:len_pdf}. Moreover, comparing the absolute proof size between $\allUipMin$ and 1-UIP yields a similar trend as clause size comparison in Fig~\ref{fig:len_compare}.

%Fig.~\ref{fig:proof_compare} shows the
%absolute proof size comparison between $\allUipMin$ and 1-UIP, and displays a similar trend %as the clause size comparison plot in Fig.~\ref{fig:len_compare}.

%\begin{figure} 
%\begin{center} 
%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
%\hline
%Solver & optimized proof size (MB) & relative reduction size  \\ 
%\hline
%$\MapleBase$ & 613.9 & 0  \\ 
%\hline
%$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
%\hline
%$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
%\hline
%\end{tabular}
%\end{center}
%\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$. %Optimized proof size measures the average absolute proof size in MB, and relative reduction %size measures the average relative reduction for all UNSAT instances.}
%\label{fig:t3}
%\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
%    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
%    \label{fig:proof_compare}
%\end{figure}

%\subsection{$\allUip$ as a Practical Learning Scheme}
%To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
%re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
%in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
%$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
%$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
%benchmark and reported solved instances, PAR-2 score and average
%clause size.

%Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
%scheme $\allUipPure$ solved the most overall instances (228) and the
%most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
%score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
%produced the shortest average clause size, but solved the second least
%instances, one more instance than the baseline 1-UIP learning. All
%configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
%solved instances, PAR-2 score and average clause size.

\subsubsection{$\allUip$ on Modern SAT solvers}
To validate $\allUip$ as practical learning schemes on modern SAT
solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
SAT Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT} and $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$). $\expSATShort$
is a top ten solver from 2019 SAT race which uses random walk
simulation to help branching. We chose $\expSATShort$ because 1) it is
a top solver in the 2019 SAT Race without using chronological
backtracking; 2) the random walk simulation 
branching heuristic is different from local branching heuristic (VSIDS and LRB) that we have considered,
hence it is in the blind spot of $\allUip$.
we compare the base 1-UIP learning scheme against $\allUipPure$,
$\allUipMin$ and the top two $\allUip$ variants, $\allUipAct$ and
$\allUipIn$, on the SAT Race 2019 main track benchmark. We report
solved instances, PAR-2 score and the average clause size.

@@

Table~\ref{fig:t5} shows the benchmark result of $\allUip$
configurations on modern solvers. Overall, we observed similar performance gain on all modern solvers as we have seen on $\MapleBase$ in Fig.~\ref{fig:t4}. More specifically, almost all configurations improved on solved instance (+4.5 instances in average) and PAR-2 score (-83.6 in average).  The average clause size reduction is consistent across all solvers. Each configuration also exhibits different strength: $\allUipPure$ solved the most instances with the best PAR-2 score on two solvers, $\allUipMin$ yields small clauses, $\allUipIn$ solved the most SAT instances, and $\allUipAct$ has stable performance.

On the SAT 2017 race winner $\MapleSeven$, all four configurations of
$\allUip$ outperformed 1-UIP learning by 12, 8, 5, and 2 instances, respectively. $\allUipPure$ solved
more UNSAT and SAT instances while other configurations improved on
solving SAT instances. The clause size reduction of $\allUip$ is more significant on $\MapleSeven$ compare to $\MapleBase$. 
%This may suggests that $\allUip$ and the recent
%learnt clause minimization approach~\cite{} synergies well because
%i-UIP clauses are shorter with more common literals which allows
%vivification~\cite{} to prune literals more aggressively through unit
%propagation.

The SAT 2018 race winner$\MapleEightShort$ uses chronological backtracking (CB) for long distance backtracks. Therefore, we applied $\allUip$-CB extension for all $\allUip$ configurations.  Three out of four $\allUip$ configurations outperformed 1-UIP by 5, 4 an 4 instances,
respectively. 
%$\MapleEightShort$ improved from $\MapleSeven$ by using
%chronological backtracking (CB) for long distance
%backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
%configurations.

%The results indicates that the improvement of $\allUip$
%is slightly shadowed by the adoption of CB. More specifically, we
%believe CB prevents decision levels from being compressed through the
%process of backtracking and literal assertion. The shorter i-UIP
%clauses can bring related literals closer, which in turns compresses
%the assertion stack and the decision levels. However, since CB
%discourages long distance backtracking, the effect of learning shorter
%clauses is shadowed until a full restart.  We believe we have observed
%an interesting interaction effect that shows the limitations of both
%CB and $\allUip$ learning for future research.

On the SAT 2019 race winner $\MapleNineShort$,  three out of four$\allUip$ configurations solved instances 6, 5 and 5 more instances than 1-UIP, respectively. The solver uses duplicate learning (DL) to prioritize clauses that are learned multiple times. We observed that $\allUip$ configurations
i-UIP clauses are less likely to be duplicated. As an example, $\allUipMin$ in average, added 12\% less duplicated clauses into the
core clause database than 1-UIP. This observation is surprising and the cause is unclear.

%One possible explanation is that
%$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
%for different assertion trails. Therefore, the solver is less likely
%to learn the same i-UIP clauses.

On $\expSATShort$,
three out of four $\allUip$ configurations solved 4, 7 and 8 more instances
than 1-UIP learning. We noticed that $\allUipAct$ and $\allUipIn$ show significant better performance than $\allUipMin$ and $\allUipPure$ on this solver. The random walk simulation branching heuristic however didn't impact the performance of $\allUip$ schemes significantly. 
%$\allUip$ to solve the similar amount of instances with close PAR-2
%scores because the additional random walk exploration allows the
%learning schemes to partially sidestep the activity problem; hence,
%the learning adjustment for variable activities should have less
%impact on the solver's performance. However, we instead, observed that
%both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
%learning scheme. One possible explanation is that $\allUipAct$ and
%$\allUipIn$ schemes could easily overcompensate variable activities
%for i-UIP clauses, and $\expSATShort$'s random walk exploration could
%use future search information to mitigate the negative effects of our
%overcompensation.
